### 2015-05-06 papers/histogram/thesis/thesis.tex

We identify and develop efficient numerical methods for determining
thermodynamic properties of the square-well fluid in order to test
square-well density functional theories. The liquid-vapor phase
interface is an interesting regime for testing density functional
theories, but unbiased Metropolis Monte Carlo simulations are
incapable of sampling the low energy fluid states which dominate the
partition function at low (i.e. liquid-state) temperatures. Previous
works have developed several generic Monte Carlo histogram methods for
collecting statistics on low energy system states, but little or no
literature exists on these methods' systematic comparison, as well as
their application to the square-well fluid. The square-well fluid in
particular introduces application challenges not manifest in
traditional models for testing and benchmarking such numerical
techniques (e.g. the Ising model). We propose to implement these
methods in Monte Carlo simulations of the square-well fluid, determine
appropriate performance metrics for their comparison, and identify
those methods which are most efficient and effective at sampling
otherwise improbable system states.

@@

\fixme{Make all mention of what we do in present, rather than past
  tense}

@@

\fixme{Give an overview of the methods section}

\fixme{Make a note on storing the logarithm of weights and
  periodically rescaling them (resetting the zero on $\ln w$) to
  reduce numerical error.}

@@

Once we identify an energy $E_-$ satisfying (\ref{eq:min_E_condition})
in an initialization routine for any histogram method, we effectively
forbid transitions to energies $E<E_-$ by tentatively setting weights
$w\p{E<E_-}=e^{-\p{E-E_-}/kT}$ for some temperature $T\ll
T_{\t{min}}$. These weights prevent the initialization routine from
wasting time on energies at which we will override the weight array
after initialization anyways. A special case was made for Wang-Landau,
which was designed assuming knowledge of the energy range of interest
a priori. In order to fairly compare Wang-Landau to other methods, we
therefore identified $E_-$ for each simulated system using other
methods, and hard-coded the minimum important energies for each system
into the Wang-Landau initialization routines\fixme{talk about how bad
  this is}. We always set $w\p{E<E_-}=e^{-\p{E-E_-}/kT}$ after
initialization of all histogram method.

@@

\subsection{Histograms and arrays}

@@

\subsection{Sample rates}

\fixme{Talk about initialization times, and how they scale with system
  size}

@@

\subsection{Errors in $u$, $c_V$, and $s_{\t{config}}$}

\fixme{Discuss errors in computed thermodynamic properties. Introduce
  Converged flat weights.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{thesis}

>>>>

We identify and develop efficient numerical methods for determining
thermodynamic properties of the square-well fluid in order to test
square-well density functional theories. The liquid-vapor phase
interface is an interesting regime for testing density functional
theories, but unbiased Metropolis Monte Carlo simulations are
incapable of sampling the low energy fluid states which dominate the
partition function at low (i.e. liquid-state) temperatures. Previous
works have developed several generic Monte Carlo histogram methods for
collecting statistics on low energy system states, but little or no
literature exists on these methods' systematic comparison, as well as
their application to the square-well fluid. The square-well fluid in
particular introduces application challenges not manifest in
traditional models for testing and benchmarking such numerical
techniques (e.g. the Ising model). We propose to implement these
methods in Monte Carlo simulations of the square-well fluid, determine
appropriate performance metrics for their comparison, and identify
those methods which are most efficient and effective at sampling
otherwise improbable system states.\fixme{make this an abstract,
  rather than a proposal...}

@@

\fixme{Always talk about what we do in our own code in present, rather
  than past tense}

@@

Having covered the basics of the square-well fluid as well as
Metropolis and biased Monte Carlo simulations, we can now discuss
several histogram methods for constructing energy weights.\fixme{more
  overview of the methods section?}

@@

Once we identify an energy $E_-$ satisfying (\ref{eq:min_E_condition})
in an initialization routine for any histogram method, we effectively
forbid transitions to energies $E<E_-$ by tentatively setting weights
$w\p{E<E_-}=e^{-\p{E-E_-}/kT}$ for some temperature $T\ll
T_{\t{min}}$. These weights prevent the initialization routine from
wasting time on energies at which we will override the weight array
after initialization anyways. A special case was made for Wang-Landau,
as the design of this method assumes knowledge of the energy range of
interest a priori. In order to fairly compare Wang-Landau to other
methods, we therefore identified $E_-$ for each simulated system using
other methods, and hard-coded the minimum important energies for each
system into the Wang-Landau initialization routines\fixme{talk about
  how bad this is}. We always set
$w\p{E<E_-}=e^{-\p{E-E_-}/kT_{\t{min}}}$ after initialization of all
histogram methods with $kT_{\t{min}}=0.2\epsilon$. We chose this
temperature because it seems to be right below the temperature at
which the square-well fluid condenses from a gas into a
liquid (see Section \ref{sec:errors}). \fixme{what about pressure?}

@@

\subsection{Histograms and arrays}
\label{sec:histograms}

@@

\subsection{Sampling rates}
\label{sec:sampling_rates}

@@

\subsection{Errors in $u$, $c_V$, and $s_{\t{config}}$}
\label{sec:errors}

\fixme{Discuss errors in computed thermodynamic properties. Introduce
  converged flat weights.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\nocite{*} \bibliography{thesis}
